---
title: "Tutorial"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(grader)

knitr::opts_chunk$set(message = FALSE)

learnr::tutorial_options(exercise.timelimit = 60,
                         exercise.checker = grader::grade_learnr)
```

## tl;dr

### Check the Code

`sqrt(log(2))`

````markdown
`r ''````{r strict, exercise = TRUE}
sqrt(log(3))
```

`r ''````{r strict-solution}
sqrt(log(2))
```

`r ''````{r strict-check}
grader::check_code("Good job. Don't worry, things will soon get harder.")
```
````

```{r strict, exercise = TRUE}
sqrt(log(3))
```

```{r strict-hint-1}
"You can take the log with `log()`"
```

```{r strict-hint-2}
"You can take the square root with `sqrt()`"
```

```{r strict-solution}
sqrt(log(2))
```

```{r strict-check}
grader::check_code("Good job. Don't worry, things will soon get harder.")
```

### Check the result

````markdown
`r ''````{r result, exercise = TRUE}
sqrt(log(3))
```

`r ''````{r result-check}
grader::check_result(
  pass_if(~ round(.result, 3) == .833, 'You got it correct!'),
  fail_if(~ round(.result, 3) == 1.048, 'You passed in 3 instead of 2'),
  correct = "wo0t!"
)
```
````

```{r result, exercise = TRUE}
sqrt(log(2 + 1))
```

```{r result-hint-1}
"You can take the log with `log()`"
```

```{r result-hint-2}
"You can take the square root with `sqrt()`"
```

```{r result-check}
grader::check_result(
  pass_if(~ round(.result, 3) == .833, 'You got it correct!'),
  fail_if(~ round(.result, 3) == 1.048, 'You passed in 3 instead of 2'),
  correct = "wo0t!"
)
```

## Billboard Rankings

Recall the 3 features of [tidy data][wickham2013]

[wickham2013]: https://vita.had.co.nz/papers/tidy-data.pdf

1. Each variable forms a column
2. Each observation forms a row
3. Each type of observational unit forms a table

Data is downloaded into the `data` folder:

```bash
wget https://raw.githubusercontent.com/hadley/tidy-data/master/data/billboard.csv
```

```{r, message=FALSE}
library(readr)
library(here)

billboard <- readr::read_csv('https://raw.githubusercontent.com/hadley/tidy-data/master/data/billboard.csv')
billboard[1:10, 1:10]
```

What's wrong with this? Well... it violates the first rule of tidy data,
the week ranking variable is spread out across columns, not in a single column.
We want a single column for the week number for a particular song,
and it's ranking for that week.

E.g.,

```
# A tibble: 10 x 10
    year artist.inverted    track                     ... week        rank
   <dbl> <chr>              <chr>                     ... <chr>       <dbl>
 1  2000 Destiny's Child    Independent Women Part I  ... x1st.week   78
 2  2000 Destiny's Child    Independent Women Part I  ... x2nd.week   63
 3  2000 Destiny's Child    Independent Women Part I  ... x2nd.week   49
```

### Tidy the data -- check_code

If we put in `spread` instead of the correct solution`, `gather`,
we will get an error about passing `spread` the wrong function parameters


```{r ebepjqxhskvqnspn, exercise = TRUE}
library(tidyr)

billboard_tidy <- spread(billboard, "week", "rank", x1st.week:x76th.week)
```

```{r ebepjqxhskvqnspn-hint-1}
# hint text
"Use the gather function from tidyr"
```

```{r ebepjqxhskvqnspn-hint-2}
# hint text
"You can use tidyselect notation to select the columns, e.g., col1:col76"
```

```{r ebepjqxhskvqnspn-solution}
# solution code
billboard_tidy <- gather(billboard, "week", "rank", x1st.week:x76th.week)
```

```{r ebepjqxhskvqnspn-check}
# check code
grader::check_code()
```

## Checking results

What if the student just finds another way?

We can just check the result, not the code.

```{r ytnsfpdawbhorgwb, exercise = TRUE}
# student code
9 / 3
```

```{r ytnsfpdawbhorgwb-check}
grader::check_result(
  grader::pass_if(~ .result == 3, "Yay!")
)
```

##### Tidy Example -- check result

Student code below does not match the solution code, but the results are the same.

```r
solution <- tidyr::gather(head(billboard), "week", "rank", x1st.week:x76th.week)
```

```{r zuzwislfkfnsoesa, exercise = TRUE}
# student code
user <- reshape2::melt(head(billboard),
                       id.vars = c("year", "artist.inverted", "track", "time",
                                   "genre", "date.entered", "date.peaked"),
                       variable.name = "week",
                       value.name = "rank",
                       factorAsStrings = TRUE)
user$week <- as.character(user$week)
```

```{r zuzwislfkfnsoesa-check}
solution <- tidyr::gather(head(billboard), "week", "rank", x1st.week:x76th.week)

grader::check_result(
  grader::pass_if(~ all.equal(tibble::as_tibble(.result), tibble::as_tibble(solution)), 'Correct!')
)
```

## Clean data

So we have a tidy dataset

```{r}
billboard_tidy <- tidyr::gather(billboard, "week", "rank", x1st.week:x76th.week)
billboard_tidy
```

But we want the week to be a numeric, not a string.

## Functions

Write a function that takes a string and extracts the numbers


### Using check result  (Works only for correct solution)

```{r}
library(stringr)
```

```{r wdyaoutxcvxxyjxa, exercise = TRUE}
# student code
my_extract_week <- function(string) {
  as.numeric(stringr::str_extract(string, '\\d+'))
}
```

```{r wdyaoutxcvxxyjxa-check}

grader::check_result(
  grader::pass_if(~ .result('x1st.week') == 1, 'hoorah'),
  grader::pass_if(~ .result('x72nd.week') == 72, 'yipee')
)
```

### Using check_code (Works only for correct solution)

```{r aoqqmwwnkvzikirk, exercise = TRUE}
# student code
my_extract_week <- function(string) {
  as.numeric(stringr::str_extract(string, '\\d+'))
}
```

```{r aoqqmwwnkvzikirk-solution}
# solution code
my_extract_week <- function(string) {
  as.numeric(stringr::str_extract(string, '\\d+'))
}
```

```{r aoqqmwwnkvzikirk-check}
# check code
grader::check_code()
```

## Apply the function

```{r}
library(stringr)
library(readr)
library(tidyr)
library(here)


extract_week <- function(string) {
  as.numeric(stringr::str_extract(string, '\\d+'))
}

billboard_tidy <- tidyr::gather(billboard, "week", "rank", x1st.week:x76th.week)
billboard_tidy
```

```{r}
library(dplyr)
library(purrr)
```

```{r otsymvcnkuaybeey, exercise = TRUE}
# student code

billboard_tidy %>%
  dplyr::mutate(week_num = purrr::map_dbl(week, extract_week))
```


```{r otsymvcnkuaybeey-solution}
# solution code
billboard_tidy %>%
  dplyr::mutate(week_num = purrr::map_dbl(week, extract_week))
```

```{r otsymvcnkuaybeey-check}
# check code
grader::check_code()
```

## Fit model

Let's go fit a model

```{r}
df <- billboard_tidy %>%
  dplyr::mutate(week_num = purrr::map_dbl(week, extract_week))
```

Can check results by comparing coefficients:

```{r ckojrogkhwnygvxp, exercise = TRUE}
# student code
fit <- lm(rank ~ week_num, data = df)
```

```
grader::check_result(
  grader::pass_if(~ all(round(broom::tidy(.result)$estimate, 2) == c(64.13, -1.14)), 'winner winner chicken dinner')
)
```

```{r ckojrogkhwnygvxp-check}
library(broom)

grader::check_result(
  grader::pass_if(~ all(round(broom::tidy(.result)$estimate, 2) == c(64.13, -1.14)), 'winner winner chicken dinner')
)
```
