---
title: "Tutorial"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(grader)

learnr::tutorial_options(#exercise.timelimit = 60,
                         exercise.checker = grader::grade_learnr)
```

## Check code examples

Here is a number. You can do great things with it, like this:

* Take the square root of the log of the number `2`.

Then click Submit Answer.

```{r strict, exercise = TRUE}
sqrt(log(2))
```

```{r strict-hint-1}
"You can take the log with `log()`"
```

```{r strict-hint-2}
"You can take the square root with `sqrt()`"
```

```{r strict-solution}
sqrt(log(2))
```

```{r strict-check}
grader::check_code("Good job. Don't worry, things will soon get harder.")
```

Using `check_result`

```{r result, exercise = TRUE}
sqrt(log(2))
```

```{r result-hint-1}
"You can take the log with `log()`"
```

```{r result-hint-2}
"You can take the square root with `sqrt()`"
```

```{r result-check}
grader::check_result(
  pass_if(~ round(.result, 3) == .833, 'You got it correct!'),
  fail_if(~ round(.result, 3) == 1.048, 'You passed in 3 instead of 2'),
  correct = "wo0t!"
)
```

## colon range

```{r range, exercise = TRUE}
1:5
```


```{r range-solution}
1:5
```

```{r range-check}
grader::check_code("Good job. Don't worry, things will soon get harder.")
```

## dplyr select

### dplyr tidyselect

```{r dplyrselect, exercise = TRUE}
library(dplyr)

mtcars %>%
  select(mpg:wt)
```


```{r dplyrselect-solution}
mtcars %>%
  select(mpg:wt)
```

```{r dplyrselect-check}
grader::check_code()
```

### dplyr tidyselect2

```{r dplyrselect2, exercise = TRUE}
library(dplyr)

mtcars %>%
  select(mpg, cyl, -wt, am:carb)
```


```{r dplyrselect2-solution}
mtcars %>%
  select(mpg, cyl, -wt, am:carb)
```

```{r dplyrselect2-check}
grader::check_code()
```

### dplyr tidyselect3

```{r dplyrselect3, exercise = TRUE}
library(dplyr)

mtcars %>%
  select(mpg, cyl, wt:carb, -am)
```


```{r dplyrselect3-solution}
mtcars %>%
  select(mpg, cyl, wt:carb, -am)
```

```{r dplyrselect3-check}
grader::check_code()
```

## tidyr


```{r tidyr1, exercise = TRUE}
library(tidyr)

who %>%
  gather(key, value, new_sp_m014:newrel_f65, na.rm = TRUE)
```


```{r tidyr1-solution}
who %>%
  gather(key, value, new_sp_m014:newrel_f65, na.rm = TRUE)
```

```{r tidyr1-check}
grader::check_code()
```


## Case Study

To finish off the chapter, let’s pull together everything you’ve learned to tackle a realistic data tidying problem. The tidyr::who dataset contains tuberculosis (TB) cases broken down by year, country, age, gender, and diagnosis method. The data comes from the 2014 World Health Organization Global Tuberculosis Report, available at http://www.who.int/tb/country/data/download/en/.

There’s a wealth of epidemiological information in this dataset, but it’s challenging to work with the data in the form that it’s provided:

```{r, message=FALSE}
library(tidyr)
library(dplyr)
```

```{r}
who
```

This is a very typical real-life example dataset. It contains redundant columns, odd variable codes, and many missing values. In short, who is messy, and we’ll need multiple steps to tidy it. Like dplyr, tidyr is designed so that each function does one thing well. That means in real-life situations you’ll usually need to string together multiple verbs into a pipeline.

The best place to start is almost always to gather together the columns that are not variables. Let’s have a look at what we’ve got:

It looks like country, iso2, and iso3 are three variables that redundantly specify the country.

year is clearly also a variable.

We don’t know what all the other columns are yet, but given the structure in the variable names (e.g. new_sp_m014, new_ep_m014, new_ep_f014) these are likely to be values, not variables.

So we need to gather together all the columns from new_sp_m014 to newrel_f65. We don’t know what those values represent yet, so we’ll give them the generic name "key". We know the cells represent the count of cases, so we’ll use the variable cases. There are a lot of missing values in the current representation, so for now we’ll use na.rm just so we can focus on the values that are present.


```{r}
who1 <- who %>% 
  tidyr::gather(new_sp_m014:newrel_f65,
                key = "key", value = "cases", na.rm = TRUE)
who1
```

We can get some hint of the structure of the values in the new key column by counting them:

```{r}
who1 %>% 
  dplyr::count(key)
```

You might be able to parse this out by yourself with a little thought and some experimentation, but luckily we have the data dictionary handy. It tells us:

The first three letters of each column denote whether the column contains new or old cases of TB. In this dataset, each column contains new cases.

The next two letters describe the type of TB:

rel stands for cases of relapse
ep stands for cases of extrapulmonary TB
sn stands for cases of pulmonary TB that could not be diagnosed by a pulmonary smear (smear negative)
sp stands for cases of pulmonary TB that could be diagnosed be a pulmonary smear (smear positive)
The sixth letter gives the sex of TB patients. The dataset groups cases by males (m) and females (f).

The remaining numbers gives the age group. The dataset groups cases into seven age groups:

014 = 0 – 14 years old
1524 = 15 – 24 years old
2534 = 25 – 34 years old
3544 = 35 – 44 years old
4554 = 45 – 54 years old
5564 = 55 – 64 years old
65 = 65 or older
We need to make a minor fix to the format of the column names: unfortunately the names are slightly inconsistent because instead of new_rel we have newrel (it’s hard to spot this here but if you don’t fix it we’ll get errors in subsequent steps). You’ll learn about str_replace() in strings, but the basic idea is pretty simple: replace the characters “newrel” with “new_rel”. This makes all variable names consistent.

```{r}
who2 <- who1 %>% 
  mutate(key = stringr::str_replace(key, "newrel", "new_rel"))
who2
```

We can separate the values in each code with two passes of separate(). The first pass will split the codes at each underscore.

```{r}
who3 <- who2 %>% 
  separate(key, c("new", "type", "sexage"), sep = "_")
who3
```

Then we might as well drop the new column because it’s constant in this dataset. While we’re dropping columns, let’s also drop iso2 and iso3 since they’re redundant.

```{r}
who3 %>% 
  count(new)
```

```{r}
who4 <- who3 %>% 
  select(-new, -iso2, -iso3)
```

Next we’ll separate sexage into sex and age by splitting after the first character:

```{r}
who5 <- who4 %>% 
  separate(sexage, c("sex", "age"), sep = 1)
who5
```

The who dataset is now tidy!

I’ve shown you the code a piece at a time, assigning each interim result to a new variable. This typically isn’t how you’d work interactively. Instead, you’d gradually build up a complex pipe:

```{r}
who %>%
  gather(key, value, new_sp_m014:newrel_f65, na.rm = TRUE) %>% 
  mutate(key = stringr::str_replace(key, "newrel", "new_rel")) %>%
  separate(key, c("new", "var", "sexage")) %>% 
  select(-new, -iso2, -iso3) %>% 
  separate(sexage, c("sex", "age"), sep = 1)
```



### Exercise 1

#### A
In this case study I set `na.rm = TRUE` just to make it easier to check that we had the correct values. Is this reasonable? Think about how missing values are represented in this dataset. Are there implicit missing values? What’s the difference between an NA and zero?


Run the `gather` the `who` dataset with `na.rm = FALSE` and count the number of rows in the dataset.

```{r gather-narm-false, exercise=TRUE}
tidyr::gather(who, key, value, ____ = ____)
```

```{r gather-narm-false-hint-1}
"Use the na.rm function argument"
```

```{r gather-narm-false-hint-2}
"Set the argument to FALSE"
```

```{r gather-narm-false-solution}
tidyr::gather(who, key, value, na.rm = FALSE)
```

```{r gather-narm-false-check}
grader::check_code("Great! Now let's see what happens when we don't drop the missing values")
```

#### B

Run the `gather` the `who` dataset with `na.rm = TRUE` and count the number of rows in the dataset.

```{r gather-narm-true, exercise=TRUE}
tidyr::gather(who, key, value, new_sp_m014:newrel_f65, na.rm = FALSE)
```

```{r gather-narm-true-solution}
tidyr::gather(who, key, value, new_sp_m014:newrel_f65, na.rm = TRUE)
```

```{r gather-narm-true-check}
grader::check_code()
```


### Exercise 2

What happens if you neglect the mutate() step?

(`mutate(key = stringr::str_replace(key, "newrel", "new_rel"))`)

```{r gather-nomutate-true, exercise=TRUE}
who %>%
  gather(key, value, new_sp_m014:newrel_f65, na.rm = TRUE) %>% 
  mutate(key = stringr::str_replace(key, "newrel", "new_rel")) %>%
  separate(key, c("new", "var", "sexage")) %>% 
  select(-new, -iso2, -iso3) %>% 
  separate(sexage, c("sex", "age"), sep = 1)
```

```{r gather-nomutate-true-solution}
who %>%
  gather(key, value, new_sp_m014:newrel_f65, na.rm = TRUE) %>% 
  separate(key, c("new", "var", "sexage")) %>% 
  select(-new, -iso2, -iso3) %>% 
  separate(sexage, c("sex", "age"), sep = 1)
```

```{r gather-nomutate-true-check}
grader::check_code()
```


### Exercise 3

I claimed that iso2 and iso3 were redundant with country. Confirm this claim.

### Exercise 4

For each country, year, and sex compute the total number of cases of TB. Make an informative visualisation of the data.

## Exercise with Code

*Here's an exercise with some prepopulated code as well as `exercise.lines = 5` to provide a bit more initial room to work.*

Now write a function that adds any two numbers and then call it:

```{r add-function, exercise=TRUE, exercise.lines = 5}
add <- function() {
  
}
```

## Topic 2

### Exercise with Hint

*Here's an exercise where the chunk is pre-evaulated via the `exercise.eval` option (so the user can see the default output we'd like them to customize). We also add a "hint" to the correct solution via the chunk immediate below labeled `print-limit-hint`.*

Modify the following code to limit the number of rows printed to 5:

```{r print-limit, exercise=TRUE, exercise.eval=TRUE}
mtcars
```

```{r print-limit-hint}
head(mtcars)
```

### Quiz

*You can include any number of single or multiple choice questions as a quiz. Use the `question` function to define a question and the `quiz` function for grouping multiple questions together.*

Some questions to verify that you understand the purposes of various base and recommended R packages:

```{r quiz}
quiz(
  question("Which package contains functions for installing other R packages?",
    answer("base"),
    answer("tools"),
    answer("utils", correct = TRUE),
    answer("codetools")
  ),
  question("Which of the R packages listed below are used to create plots?",
    answer("lattice", correct = TRUE),
    answer("tools"),
    answer("stats"),
    answer("grid", correct = TRUE)
  )
)
```
